---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metaLik2

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![R-CMD-check](https://github.com/giuseppealfonzetti/metaLik2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/giuseppealfonzetti/metaLik2/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `metaLik2` provides higher order likelihood inference for random-effects
meta-analysis and meta-regression models. It leverages the `likelihoodAsy` package to provide a modular extension to further models than 

## Installation

You can install the development version of metaLik2 like so:

``` r
pak::pkg_install("giuseppealfonzetti/metaLik2")
```

## Continuous outcomes

We fit the same random-effects model to the `diuretics` data shipped with
`metaLik`, using both `metaLik()` and `metaLik2()`

```{r setup}
library(metaLik)
library(metaLik2)
data(diuretics)

dN <- metaLik2_set_data(diuretics, "continuous")
dN
fitNN <- metaLik2::metaLik2("normal_normal", dN)
fitNN0 <- metaLik::metaLik(y ~ 1, data = diuretics, sigma2 = sigma2)
```

### A metaLik2 fit

The fit object supports `print()`, `coef()`, `vcov()` and `logLik()` methods.

```{r methods}
fitNN 
coef(fitNN)
vcov(fitNN)
logLik(fitNN)

coef(fitNN0)
log(fitNN0$tau2.mle)
logLik(fitNN0) # constant term dropped
```

The `confint()` method provides Wald confidence intervals

```{r}
confint(fitNN)
```

### Hypothesis testing

Both packages test a scalar parameter using the
`r` statistic or Skovgaard's `r*`. In `metaLik2()` computations are run via `likelihoodAsy`

```{r test}
metaLik2::rstar_test(fitNN, PARAM = 1, ALTERNATIVE = "greater")
metaLik::test.metaLik(fitNN0, param = 1, alternative = "greater")
```

### Confidence intervals

Confidence intervals can be computed via `rstar_ci()`, which returns an `rstarci` object from the `likelihoodAsy` package

```{r rstarci}
ci <- metaLik2::rstar_ci(fitNN, SEED = 123)
class(ci)
```

which comes with `print()`, `summary()` and `plot()` methods.

```{r}
ci
summary(ci)
plot(ci)
```

## Binary outcomes

For studies reporting a 2x2 table of event counts, we use the `dat.bcg`
BCG-vaccine data shipped with `metafor`, reshaped into the
`event1/n1/event2/n2` layout that `metaLik2_set_data(..., "binary")` expects.

```{r binary-data, message=FALSE}
library(metafor)
counts <- with(dat.bcg, data.frame(
  event1 = tpos, n1 = tpos + tneg,
  event2 = cpos, n2 = cpos + cneg
))
dB <- metaLik2_set_data(counts, "binary")
head(counts)
```

### Binomial-normal

The `"binomial_normal"` model fits a binomial-within-study,
normal-between-study random-effects model. The overall log-odds ratio is the
`delta` parameter.

```{r binary-normal-fit}
fitBN <- metaLik2("binomial_normal", dB)
c(delta = coef(fitBN)[["delta"]], tau2 = exp(coef(fitBN)[["log_tau2"]]))
```

The same model is available in `metafor` as a mixed-effects logistic
regression, and the Wald confidence intervals coincide.

```{r binary-normal-metafor}
fitBN0 <- rma.glmm(measure = "OR", ai = tpos, bi = tneg, ci = cpos, di = cneg,
                   data = dat.bcg, model = "UM.FS")
c(delta = as.numeric(fitBN0$beta), tau2 = fitBN0$tau2)
c(summary(fitBN0)$ci.lb, summary(fitBN0)$ci.ub)
confint(fitBN)
```

Higher-order inference on `delta` uses the same `r*` interface as the
continuous case.

```{r binary-normal-inference, warning=FALSE}
metaLik2::rstar_test(fitBN, PARAM = "delta", R = 200)
metaLik2::rstar_ci(fitBN, PARAM = "delta", R = 200)
```

### Binomial-beta

The `"binomial_beta"` model instead gives each arm a beta-binomial marginal and
couples the two arms through a Sarmanov dependence term, so both arm
probabilities are random. Its parameters are the log beta shapes per arm
(`log_a1`, `log_b1`, `log_a2`, `log_b2`) plus a dependence parameter `eta`.

```{r binary-beta-fit}
fitBB <- metaLik2("binomial_beta", dB)
coef(fitBB)
```

Here the scalar of interest `delta = log(a1 / b1) - log(a2 / b2)` is the
log-odds ratio contrast between the two arm means, and takes the same `r*`
interface. 

```{r binary-beta-inference, warning=FALSE}
metaLik2::rstar_test(fitBB, PARAM = "delta", R = 200)
```