---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metaLik2

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![R-CMD-check](https://github.com/giuseppealfonzetti/metaLik2/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/giuseppealfonzetti/metaLik2/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `metaLik2` provides higher order likelihood inference for random-effects
meta-analysis and meta-regression models. It leverages the `likelihoodAsy` package to provide a modular extension to further models than 

## Installation

You can install the development version of metaLik2 like so:

``` r
pak::pkg_install("giuseppealfonzetti/metaLik2")
```

## Usage

We fit the same random-effects model to the `diuretics` data shipped with
`metaLik`, using both `metaLik()` and `metaLik2()`

```{r setup}
library(metaLik)
library(metaLik2)
data(diuretics)

d <- metaLik2_set_data(diuretics, "continuous")
d
fit <- metaLik2::metaLik2("normal_normal", d)
fit0 <- metaLik::metaLik(y ~ 1, data = diuretics, sigma2 = sigma2)
```

### A metaLik2 fit

The fit object supports `print()`, `coef()`, `vcov()` and `logLik()` methods.

```{r methods}
fit 
coef(fit)
vcov(fit)
logLik(fit)

coef(fit0)
log(fit0$tau2.mle)
logLik(fit0) # constant term dropped
```

The `confint()` method provides Wald confidence intervals

```{r}
confint(fit)
```

### Hypothesis testing

Both packages test a scalar parameter using the
`r` statistic or Skovgaard's `r*`. In `metaLik2()` computations are run via `likelihoodAsy`

```{r test}
metaLik2::rstar_test(fit, PARAM = 1, ALTERNATIVE = "greater")
metaLik::test.metaLik(fit0, param = 1, alternative = "greater")
```

### Confidence intervals

Confidence intervals can be computed via `rstar_ci()`, which returns an `rstarci` object from the `likelihoodAsy` package

```{r rstarci}
pr <- metaLik2::rstar_ci(fit, SEED = 123)
class(pr)
```

which comes with `print()`, `summary()` and `plot()` methods.

```{r}
pr
summary(pr)
plot(pr)
```

## Binary outcomes


```{r binary-data, message=FALSE}
library(metafor)
counts <- with(dat.bcg, data.frame(
  event1 = tpos, n1 = tpos + tneg,
  event2 = cpos, n2 = cpos + cneg
))
head(counts)
```

We fit a `"binomial_normal"` using `metaLik2()`. The overall
log-odds ratio is the `delta` parameter.

```{r binary-fit}
db <- metaLik2_set_data(counts, "binary")
fitb <- metaLik2("binomial_normal", db)
c(delta = coef(fitb)[["delta"]], tau2 = exp(coef(fitb)[["log_tau2"]]))
```

The same model is available in `metafor` as a mixed-effects logistic regression.

```{r binary-metafor}
fitb0 <- rma.glmm(measure = "OR", ai = tpos, bi = tneg, ci = cpos, di = cneg,
              data = dat.bcg, model = "UM.FS")
c(delta = as.numeric(fitb0$beta), tau2 = fitb0$tau2)
```

The `metafor` package returns Wald confidence intervals, which coincide with `metaLik2` confint()

```{r}
c(summary(fitb0)$ci.lb,summary(fitb0)$ci.ub)
confint(fitb)
```


Higher-order inference on `delta` can be run using the same `r*` interface as the continuous
case

```{r binary-inference, warning=FALSE}
metaLik2::rstar_test(fitb, PARAM = "delta", R = 200)
metaLik2::rstar_ci(fitb, PARAM = "delta", R = 200)
```